---
title: "Finding Novel Exons by Princess Kuteyi"
format: html
editor: visual
---

## Finding Novel Exons

The beginning portion of this script is to load the packages needed and also loading the data for the analysis.

```{r}
library(GenomicRanges)
library(rtracklayer)
library(Repitools)
library(DESeq2)

```

## Downloading BiocManager and Recount3

Using Bioconductor and Recount3 is essential for this project so the first step is to download them.

```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")
BiocManager::install("recount3")
BiocManager::install("snapcount")
BiocManager::install("Repitools")
library("recount3")
library("snapcount")
```

Find available projects to utilize. The focus is human samples.

```{r}
human_projects <- available_projects()
dim(human_projects)
head(human_projects)
```

Pick the project and sample to analyze.

```{r}
# subset the project of interest
 potential_info <- subset(
 human_projects,
   project == "SRP107565" & project_type == "data_sources"
)
 
potential_info

```

```{r}
# utilize gtext as the file source
subset(human_projects, file_source == "gtex" & project_type == "data_sources")
```

```{r}
rse_gene_SRP107565 <- create_rse(potential_info)
rse_gene_SRP107565
metadata(rse_gene_SRP107565)

```

```{r}
dim(rse_gene_SRP107565)
rowRanges(rse_gene_SRP107565)
```
